Behavior Driven Development in Haskell

### 2012-05-24 paper.tex
brushing up intro.

Why do we dislike to write test code?
It is probably because writing test code is not fun.
To encourage Haskellers to write more test code
to detect value errors,
we need a scheme to make writing test code more interesting.

\begin{quote}
    \emph{We should explain what is BDD around here}
\end{quote}

For this purpose,
we brought design methods from other programming language communities and
integrated them with a Haskell-original test method.
In particular, we introduced {\tt doctest}~\cite{doctest} and
integrated with QuickCheck~\cite{quickcheck}. 
We also implemented {\tt hspec} inspired by rspec~\cite{rspec} 
and integrated with QuickCheck.

In our method, design specification takes role of documentation for users and
is used as test cases by its developer. 
This method would motivate Haskellers
bacause it has a kill-three-birds-with-one-stone aspect.

%The purpose of this paper is to encourage Haskllers to write documentation, which can be used for testing, in their libraries.

%%%% State your contributions.

@@

@@

\begin{description}
\item[unit test] tests for individual programs or libraries.
\item[example] Typical usage of functions.
\item[property] Invariants which functions have.
\item[behavior] How do functions behave for users. Behavior can be shown by using both examples and properties.
\item[design] TBD
\item[documentation] Explanation of functions.
\end{description}

@@

@@

\noindent Note that the placeholders library provide better keyword such as {\tt notImplemented} and {\tt todo}. Next we wrote documentation.

@@

\begin{figure*}
\small
\begin{verbatim}
  describe "rev" $ do
    it "return the first element in the last" $ property $
        \xs -> not (null xs) ==> head (rev xs) == last (xs :: [Int])

    it "return the last element in the first" $ property $
        \xs -> not (null xs) ==> last (rev xs) == head (xs :: [Int])
\end{verbatim}
\end{figure*}

\noindent We should run test in order to find errors in Haddock annotation.
Then we implement the function.

rev :: [a] -> [a]
rev = foldl (flip (:)) []
\end{verbatim}
\end{quote}

\noindent And test.

The authors are not fundamentalist.
If functions to be defined are simple enough,
we might implement them first withtout signatures.
And their signatures can be automatically
inserted by IDE (like ghc-mod) later
if Haskell compilers can infer correctly.
Also, we can cut and paste typical examples and their results
from Haskell interpreters to their document.

>>>>

In other programming language communities, 
behavior driven development (BDD) is getting much attention
slowly replacing test driven development (TDD).
In BDD, test code is written with specification terminology.
In other words, specification can be used for automatic testing.

In this paper, we consider how put BDD into practice in Haskell.
We brought BDD methods from other programming language communities and
integrated them with a Haskell-original test method.
In particular, we introduced {\tt doctest}~\cite{doctest} and
integrated with QuickCheck~\cite{quickcheck}. 
We also implemented {\tt hspec} inspired by rspec~\cite{rspec} 
and integrated with QuickCheck.

@@

\noindent Our {\tt doctest} implementation 
can be used for design, documentation and automatic testing.
We hope that this kill-three-birds-with-one-stone aspect 
motives Haskellers to write test code.

@@

\begin{description}
\item[developer] programmers who develop libraries
\item[users] programmers who use libraries
\item[unit test] tests for individual programs or libraries.
\item[example] Typical usage of functions.
\item[property] Invariants which functions have.
\item[behavior] How do functions behave for users. Behavior can be shown by using both examples and properties.
\item[design] TBD
\item[documentation] Explanation of functions.
\end{description}

@@

This section show best current practice of
BDD in Haskell.

\subsection{Procedure of BBD}

@@

\noindent Note that the placeholders library provide better keyword such as {\tt notImplemented} and {\tt todo}. 
Next we wrote visible behaviors in Haddock annotation:

@@

\noindent We should execute {\tt doctest} to ensure no syntax errors.
Also, we specifies internal behaviors with {\tt hspec}:

\begin{quote}
\small
\begin{verbatim}
describe "rev" $ do
  it "returns the first element in the last" $ 
    property $ \xs -> not (null xs) ==>
      head (rev xs) == last (xs :: [Int])

  it "returns the last element in the first" $ 
    property $ \xs -> not (null xs) ==>
      last (rev xs) == head (xs :: [Int])
\end{verbatim}
\end{quote}

\noindent Also, we should execute this hspec module
to ensure no syntax errors.
Then we implement the function.

\begin{quote}
\small
\begin{verbatim}
rev :: [a] -> [a]
rev = foldl (flip (:)) []
\end{verbatim}
\end{quote}

\noindent Now we run test suites until
all test failures disappear.

Note that the authors are not fundamentalist.
If functions to be defined are simple enough,
we might implement them first withtout signatures.
And their signatures can be automatically
inserted by IDE (like ghc-mod) later
if Haskell compilers can infer correctly.
Also, we can cut and paste examples and their results
from Haskell interpreters to their Haddock annotations.
